Fast Regression of the Tritium Breeding Ratio in Tokamak Fusion Reactors

### 2020-06-14 paper/conclusion.tex

We employed a broad spectrum of data
analysis and machine learning techniques to develop fast and high-quality
surrogates for the Paramak TBR model. After reviewing
nine surrogate model families, examining their behaviour on a constrained and
unrestricted feature space, and studying their scaling properties, we retrained
the best-performing instances to produce properties desirable for
practical use. The fastest surrogate, an artificial neural network trained
on~\num{500000} datapoints, attained an~$R^2=\num{0.985}$ with mean prediction time
of~$\SI{0.898}{\micro\second}$, representing a relative
speedup of $8\cdot 10^6$ with respect to the MC model. We demonstrated the possibility of achieving comparable results using only a
training set of size~\num{10000}.

@@

>>>>

We employed a broad spectrum of data
analysis and machine learning techniques to develop fast and high-quality
surrogates for the Paramak TBR model. After reviewing
9~surrogate model families, examining their behaviour on a constrained and
unrestricted feature space, and studying their scaling properties, we retrained
the best-performing instances to produce properties desirable for
practical use. The fastest surrogate, an artificial neural network trained
on~\num{500000} datapoints, attained an~$R^2=\num{0.985}$ with mean prediction time
of~$\SI{0.898}{\micro\second}$, representing a relative
speedup of $8\cdot 10^6$ with respect to the MC model. We demonstrated the possibility of achieving comparable results using only a
training set of size~\num{10000}.

@@

Relevant source code, model instances and datasets are freely
available online as well as a more detailed technical
report~\cite{github,finalreport}.

### 2020-06-14 paper/methodology.tex

\begin{table*}[t]
	\caption{\label{tbl:metrics}Metrics recorded in experiments. In
	formulations, we work with a training set of size $N_0$ and a test set of
size $N$, values $y^{(i)}=f(x^{(i)})$ and $\hat{y}^{(i)}=\hat{f}(x^{(i)})$
denote images of the $i$th testing sample in the MC TBR model and the surrogate
respectively. The mean $\overline{y}=N^{-1}\sum_{i=1}^N y^{(i)}$ and $P$ is the
number of input features.}
	\begin{indented}
	\item[]
		\begin{tabular}{lrl}
		\toprule
		Regression performance metrics& Notation	& Mathematical formulation\\
		\midrule
		Mean absolute error	& MAE & $N^{-1}\sum_{i=1}^N |y^{(i)}-\hat{y}^{(i)}|$ \\
		Standard error of regression & $S$	& $\text{StdDev}_{i=1}^N\left\{ |y^{(i)} -
		\hat{y}^{(i)}| \right\} $ \\
			Coefficient of determination & $R^2$	& $1-\sum_{i=1}^N
			\left(y^{(i)}-\hat{y}^{(i)} \right)^2\left[\sum_{i=1}^N \left(
			y^{(i)}-\overline{y} \right)^2\right]^{-1} $ \\
			Adjusted $R^2$ & $R^2_\text{adj.}$	& $1-(1-R^2)(N-1)(N-P-1)^{-1}$ \\
		\midrule
		Computational complexity metrics	& {}	& {} \\
		\midrule
		Mean training time & $\overline{t}_{\text{trn.}}$	& $(\text{wall training time of
		$\hat{f}(x)$})N_0^{-1}$  \\
			Mean prediction time & $\overline{t}_{\text{pred.}}$	& $(\text{wall prediction time of
			$\hat{f}(x)$})N^{-1}$ \\
				Relative speedup & $\omega$	& $(\text{wall evaluation time of $f(x)$})
				(N\overline{t}_{\text{pred.}})^{-1}$ \\
		\bottomrule
		\end{tabular}
	\end{indented}
\end{table*}

>>>>

\begin{table*}[t]
	\caption{\label{tbl:metrics}Metrics recorded in experiments. In
	formulations, we work with a training set of size $N_0$ and a test set of
size $N$, values $y^{(i)}=f(x^{(i)})$ and $\hat{y}^{(i)}=\hat{f}(x^{(i)})$
denote images of the $i$th testing sample in the MC TBR model and the surrogate
respectively. The mean $\overline{y}=N^{-1}\sum_{i=1}^N y^{(i)}$ and $P$ is the
number of input features.}
	\begin{indented}
	\item[]
		\begin{tabularx}{\textwidth}{Xrl}
		\toprule
		Regression performance metrics& Notation	& Mathematical formulation\\
		\midrule
		Mean absolute error	& MAE & $N^{-1}\sum_{i=1}^N |y^{(i)}-\hat{y}^{(i)}|$ \\
		Standard deviation of error & $S$	& $\text{StdDev}_{i=1}^N\left\{ |y^{(i)} -
		\hat{y}^{(i)}| \right\} $ \\
			Coefficient of determination & $R^2$	& $1-\sum_{i=1}^N
			\left(y^{(i)}-\hat{y}^{(i)} \right)^2\left[\sum_{i=1}^N \left(
			y^{(i)}-\overline{y} \right)^2\right]^{-1} $ \\
			Adjusted $R^2$ & $R^2_\text{adj.}$	& $1-(1-R^2)(N-1)(N-P-1)^{-1}$ \\
		\midrule
		Computational complexity metrics	& {}	& {} \\
		\midrule
		Mean training time & $\overline{t}_{\text{trn.}}$	& $(\text{wall training time of
		$\hat{f}(x)$})N_0^{-1}$  \\
			Mean prediction time & $\overline{t}_{\text{pred.}}$	& $(\text{wall prediction time of
			$\hat{f}(x)$})N^{-1}$ \\
				Relative speedup & $\omega$	& $(\text{wall evaluation$^\text{b}$ time of $f(x)$})
				(N\overline{t}_{\text{pred.}})^{-1}$ \\
		\bottomrule
		\end{tabularx}\\%
		{\footnotesize $^\text{b}$This corresponds to evaluation of the
		expensive MC TBR model on all points of the test set. In surrogates, the
		same time period is referred to as ``prediction time.''}
	\end{indented}
\end{table*}

### 2020-06-14 paper/results.tex

\subsubsection{Model Comparison}
In Experiment~4, we aim to create models that yield:
(a)~the best regression performance regardless
of other features, (b)~acceptable performance with the shortest mean
prediction time, or (c)~acceptable performance with the smallest training set.
To this end, we trained 8~surrogates that are presented in~\Fref{fig:reg-performance}
and~\Tref{tbl:exp4-detailed-results}.
Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
Experiments~2 \&~3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (Model~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline. Satisfying
goal~(b), the fastest model (Model~2, ANN) achieved~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy
goal~(c) we also trained a more simplified model (Model~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig7_qasssampling.pdf}
	\caption{\label{fig:qasssampling}Absolute training error for QASS, baseline scheme, and mixed scheme.}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig8_qasssampling100k.pdf}
	\caption{\label{fig:qasssampling100k}Absolute training error for QASS and baseline scheme, with 100k initial samples.}
\end{figure}

>>>>

\subsubsection{Model Comparison}
In Experiment~4, we aim to create models that yield:
(a)~the best regression performance regardless
of other features, (b)~acceptable performance with the shortest mean
prediction time, or (c)~acceptable performance with the smallest training set.
To this end, we trained 8~surrogates that are presented in~\Fref{fig:reg-performance}
and~\Tref{tbl:exp4-detailed-results}.
Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
Experiments~2 \&~3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (Model~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard deviation of error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline. Satisfying
goal~(b), the fastest model (Model~2, ANN) achieved~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy
goal~(c) we also trained a more simplified model (Model~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

@@

\begin{figure}
	\centering
	\hspace*{-1.5em}\includegraphics[width=1.15\linewidth]{fig7_qasssampling.pdf}
	\caption{\label{fig:qasssampling}Absolute training error for QASS, baseline scheme, and mixed scheme.}
\end{figure}

\begin{figure}
	\centering
	\hspace*{-1.5em}\includegraphics[width=1.15\linewidth]{fig8_qasssampling100k.pdf}
	\caption{\label{fig:qasssampling100k}Absolute training error for QASS and baseline scheme, with 100k initial samples.}
\end{figure}

### 2020-06-14 paper/ucl_tbr_paper.tex
correct author name
added figure and changed reactor type
rasterize figure 6 plots (reducing the paper PDF size 10x), add rasterization script
more graphics tweaks

\usepackage{siunitx}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % TODO: remove me before submitting
\usepackage{isotope}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{url}
\usepackage{graphicx}

@@

\graphicspath{{./fig/}}

@@

\author{P~Mánek$^{1,2}$, G~Van Goffrier$^1$, V~Gopakumar$^3$, N~Nikolau$^1$, J~Shimwell$^3$ and I~Waldmann$^1$}

@@

% TODO: revise these
\keywords{machine learning, surrogate model, nuclear fusion, tritium breeding, adaptive sampling}
\submitto{\NF}
\maketitle
\ioptwocol

>>>>

\usepackage{siunitx}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % TODO: remove me before submitting
\usepackage{isotope}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{url}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tabularx}

@@

\graphicspath{{./fig_final_report/}{./fig/}}

@@

\author{P~Mánek$^{1,2}$, G~Van Goffrier$^1$, V~Gopakumar$^3$, N~Nikolaou$^1$, J~Shimwell$^3$ and I~Waldmann$^1$}

@@

% TODO: revise these
\keywords{machine learning, surrogate model, regression, Monte Carlo simulation, tritium breeding, adaptive sampling}
\submitto{\NF}
\maketitle
\ioptwocol
